---
title: "Data Analysis : Assignment 2"
author: "Ghazal Ayobi and Shah Ali Gardezi"
output: pdf_document
---


```{r include=FALSE}
# CLEAR MEMORY
rm(list=ls())


# Import libraries
library(tidyverse)
library(haven)
library(data.table)
library(rms)
library(lspline)
library(huxtable)
library(modelsummary)
library(pscl)


```

```{r include=FALSE}
# Loading the data
# Get the dataset: now we use all the observations from Europe:

hotels_europe_price <- read_csv("https://osf.io/p6tyr/download")
hotels_europe_features <- read_csv("https://osf.io/utwjs/download")
```

```{r include=FALSE}
# Joining Price and Features

# Join them by hotel_id
data <- left_join(hotels_europe_price, hotels_europe_features, by = "hotel_id")
rm(hotels_europe_price,hotels_europe_features)
```


```{r include=FALSE}
# Selecting a City
data <- data %>% filter(city_actual=="Paris")

#unique(data$accommodation_type)
data <- data <- data[data$accommodation_type=="Hotel",]
 

data <- data %>% filter(price<=600) %>% filter(!is.na(stars)) %>% filter(!is.na(distance)) %>% filter(!is.na(rating))


# LOG transformation of Price
data$lnprice <- log(data$price)


data$highly_rated <- ifelse(data$rating>=4, 1, 0)
#data$distance2 <- ifelse(data$distance<=2, 1, 0)

# check for duplicates
#data<-data[!duplicated(data),]
```

## Introduction

The Question of this case study is how hotels distance from center is related to highly rated hotels. For this assignment we use Hotels-Europe data. This data set contains two tables **[Features](https://osf.io/utwjs/download)** and **[Price](https://osf.io/p6tyr/download)**. We joined the two tables using left join. 

## Data Transformation
 
As a process for filtering and data transformation, we use hotel user rating as the dependent variables and transformed it to a binary variable called highly_rated=1 if rating ≥ 4, 0 otherwise. We selected _**Paris**_ City and considered _**Hotels**_ as accommodation type. Moreover, we excluded hotels with less than _USD 600_ per night. Removed NULL and duplicated values from the data set. 

## Analysis

In the first place in order to understand what functional form to include in the regression. e use dsitacne from city center and number pf stars relationship. To caputre nonliterary 




```{r echo=FALSE}
# 2) Quick reminder: check the descriptives + association
# Summary statistics on price and log of price
P95 <- function(x){ quantile(x,.95,na.rm=T)}
datasummary( highly_rated + distance + stars ~ Mean + SD + Min + Max + Median + P95 + N , data = data )


```


- Filtered the data for **Paris**
- Selected Hotels and Apartments for accommodation type
- Price is less than _*600*_
- Removed null values from _stars_
- Removed duplicates
- Removed Null values from _rating_, _stars_, and _distance_
- Created log of price, _lnprice_
- Created _highly_rated_ if rating >= 4
- Distance is with in 2 miles


```{r include=FALSE}

# lpm (repeating the previous regression)
lpm <-lm(highly_rated ~ lspline(stars, c(4)) + distance, data=data)
summary(lpm, vcov=sandwich)

data$pred_lpm <- predict(lpm)
summary(data$pred_lpm)

```

```{r include=FALSE}
# logit coefficients
library(lspline)

logit <- glm(highly_rated ~ lspline(stars, c(4)) + distance, data=data, family='binomial')
summary(logit)
glance(logit)
#TODO
# consider change to margins package


# predicted probabilities 
data$pred_logit <- predict.glm(logit, type="response")
summary(data$pred_logit)


# logit marginal differences
library(mfx)
logit_marg <- logitmfx(formula = highly_rated ~ lspline(stars, c(4)) + distance, data=data, atmean=FALSE)

```


```{r include=FALSE}

# probit coefficients
probit <- glm(highly_rated ~ lspline(stars, c(4)) + distance, data=data, family=binomial(link="probit"))


# predicted probabilities 
data$pred_probit<- predict.glm(probit, type="response") 
summary(data$pred_probit)

# probit marginal differences
probit_marg <- probitmfx(formula = highly_rated ~ lspline(stars, c(4)) + distance, data=data, atmean=FALSE)

```



```{r echo=FALSE}
cm <- c('(Intercept)' = 'Constant')
msummary(list(lpm, logit, logit_marg, probit, probit_marg),
         fmt="%.3f",
         gof_omit = 'DF|Deviance|Log.Lik.|F|R2 Adj.|AIC|BIC|R2|PseudoR2',
         stars=c('*' = .05, '**' = .01),
         coef_rename = cm,
         coef_omit = 'as.factor(country)*'
         
)
```


```{r echo=FALSE}
# adding pseudo R2 (not work for mfx)
glance_custom.glm <- function(x) data.frame(`PseudoR2` = pR2(x)["McFadden"])
cm <- c('(Intercept)' = 'Constant')
msummary(list(lpm, logit, probit),
         fmt="%.3f",
         gof_omit = 'DF|Deviance|Log.Lik.|F|R2 Adj.|AIC|BIC',
         stars=c('*' = .05, '**' = .01),
         coef_rename = cm
)
```


```{r echo=FALSE}
g5 <- ggplot(data = data) +
  geom_point(aes(x=pred_lpm, y=pred_probit, color="Probit"), size=0.5,  shape=16) +
  geom_point(aes(x=pred_lpm, y=pred_logit,  color="Logit"), size=0.5,  shape=16) +
  geom_line(aes(x=pred_lpm, y=pred_lpm,    color="45 Degree line"), size=0.5) +
  labs(x = "Predicted probability of Highly Rated (LPM)", y="Predicted probability")+
  scale_y_continuous(expand = c(0.00,0.0), limits = c(0,1), breaks = seq(0,1,0.2)) +
  scale_x_continuous(expand = c(0.00,0.0), limits = c(0,1), breaks = seq(0,1,0.2)) +
  scale_color_manual(name = "", values=c("#541352FF", "#3a5e8cFF","#10a53dFF")) +
  theme_light()+
  theme(legend.position=c(0.55,0.08),
        legend.direction = "horizontal",
        legend.text = element_text(size = 7))
g5
```

```{r}
#distance
g1 <- ggplot(data = data, aes(x=distance, y=highly_rated)) +
  geom_smooth(method="loess", color="3a5e8cFF") +
  scale_y_continuous(expand = c(0.01,0.01),limits = c(0,1), breaks = seq(0,1,0.2), labels = scales::percent) +
  labs(x = "Distance",y = "Probability of Highly Rated") +
  theme_light() 
g1
```

```{r}
#stars

g1 <- ggplot(data = data, aes(x=stars, y=highly_rated)) +
  geom_smooth(method="loess", color="3a5e8cFF") +
  scale_y_continuous(expand = c(0.01,0.01),limits = c(0,1), breaks = seq(0,1,0.2), labels = scales::percent) +
  labs(x = "Stars",y = "Probability of Highly Rated") +
  theme_light() 
g1

```





