---
title: "Data Analysis : Assignment 2"
author: "Ghazal Ayobi and Shah Ali Gardezi"
output: pdf_document
---

```{r include=FALSE, message=FALSE, warning=FALSE}
# CLEAR MEMORY
rm(list=ls())


# Import libraries
library(tidyverse)
library(haven)
library(data.table)
library(rms)
library(lspline)
library(huxtable)
library(modelsummary)
library(pscl)
library(esquisse)
library(ggplot2)


```



```{r include=FALSE, message=FALSE, warning=FALSE}
# Loading the data
# Get the dataset: now we use all the observations from Europe:

hotels_europe_price <- read_csv("https://osf.io/p6tyr/download")
hotels_europe_features <- read_csv("https://osf.io/utwjs/download")
```

```{r include=FALSE, message=FALSE, warning=FALSE}
# Joining Price and Features

# Join them by hotel_id
data <- left_join(hotels_europe_price, hotels_europe_features, by = "hotel_id")
rm(hotels_europe_price,hotels_europe_features)
```


```{r include=FALSE, message=FALSE, warning=FALSE}
# Selecting a City
data <- data %>% filter(city_actual=="Paris")

#unique(data$accommodation_type)
data <- data <- data[data$accommodation_type=="Hotel",]
 

data <- data %>% filter(price<=600) %>% filter(!is.na(stars)) %>% filter(!is.na(distance)) %>% filter(!is.na(rating))


# LOG transformation of Price
data$lnprice <- log(data$price)


data$highly_rated <- ifelse(data$rating>=4, 1, 0)
data$high_stars <- ifelse(data$stars>=3, 1, 0)

# check for duplicates
#data<-data[!duplicated(data),]
```

## Introduction

The Question of this case study is how hotels distance from center is related to highly rated hotels. For this assignment we use Hotels-Europe data. This data set contains two tables [Features](https://osf.io/utwjs/download) and [Price](https://osf.io/p6tyr/download). We joined the two tables using left join. 


## Data Transformation

As a process of filtering and data transformation, we use hotel user rating as the dependent variables and transformed it to a binary variable called _highly_rated_ which equals to one if _rating_ is more than _4_, _0_ otherwise.  We selected _**Paris**_ City and considered _**Hotels**_ as accommodation type. Moreover, we excluded hotels with less than _USD600_ per night, removed null and duplicated values from the data set. 

## Analysis

In the first place in order to understand what functional form to include in the regression. We examined lowess regression with highly rated hotels to decide in what functional form to include stars and distance. The figures in the appendix shows that distance and highly rated hotels have negative relationship. distance increases by one mile and. between 1.2 miles and 3 miles highly rated hotels and distance does not indicate in correlation. after 3 miles, highly rated declines. as the graph indicates after 3 stars the relationship between highly rated and stars changes. Price and highly rated has interesting relationship one unit change in price till USD 250 has positive relationship, after it declines. Thus, we included distance as a piece-wise linear spline with knots at 1.2 and 3 miles and stars with a knot at 3 stars. For additional variables we included log of price as a piece-wise linear spline with knot at 5.5. Other variables are _price_ which was transformed to log of Price and _weekend_. 






```{r echo=FALSE, message=FALSE, warning=FALSE}

# Summary statistics on price and log of price

P95 <- function(x){ quantile(x,.95,na.rm=T)}
datasummary( highly_rated + distance + stars ~ Mean + SD + Min + Max + Median + P95 + N , data = data, title = "Summary Statistics" )


```


- Filtered the data for **Paris**
- Selected Hotels and Apartments for accommodation type
- Price is less than _*600*_
- Removed null values from _stars_
- Removed duplicates
- Removed Null values from _rating_, _stars_, and _distance_
- Created log of price, _lnprice_
- Created _highly_rated_ if rating >= 4

```{r include=FALSE, message=FALSE, warning=FALSE}
model_formula <- formula(highly_rated ~ high_stars + lspline(distance, c(1.2, 3)) + price + weekend )
```


```{r include=FALSE, message=FALSE, warning=FALSE}

# lpm (repeating the previous regression)
lpm <-lm(model_formula, data=data, vcov = "hetreo")
summary(lpm, vcov=sandwich)

data$pred_lpm <- predict(lpm)
summary(data$pred_lpm)

```


```{r include=FALSE, message=FALSE, warning=FALSE}
# Logit coefficients

logit <- glm(model_formula, data=data, family = binomial ( link = "logit"))

# Predicted probabilities 
data$pred_logit <- predict.glm(logit, type="response")


# Logit marginal differences
library(mfx)
logit_marg <- logitmfx(model_formula, data=data, atmean=FALSE, robust = T)

```


```{r include=FALSE, message=FALSE, warning=FALSE}

# Probit coefficients
probit <- glm(model_formula, data=data, family=binomial(link="probit"))


# Predicted probabilities 
data$pred_probit<- predict.glm(probit, type="response") 

# Probit marginal differences
probit_marg <- probitmfx(model_formula, data=data, atmean=FALSE, robust = T)

```



```{r}
cm <- c('(Intercept)' = 'Constant')
summary1 <- msummary(list(lpm, logit, logit_marg, probit, probit_marg),
         fmt="%.3f",
         gof_omit = 'DF|Deviance|Log.Lik.|F|R2 Adj.|AIC|BIC|R2|PseudoR2',
         stars=c('*' = .05, '**' = .01),
         coef_rename = cm,
         coef_omit = 'as.factor(country)*'
         
)
```



```{r}
# adding pseudo R2 (not work for mfx)
glance_custom.glm <- function(x) data.frame(`PseudoR2` = pR2(x)["McFadden"])
cm <- c('(Intercept)' = 'Constant')

summary2 <- msummary(list(lpm, logit, probit),
         fmt="%.3f",
         gof_omit = 'DF|Deviance|Log.Lik.|F|R2 Adj.|AIC|BIC',
         stars=c('*' = .05, '**' = .01),
         coef_rename = cm
)
```



```{r message=FALSE, warning=FALSE}

#distance
g1 <- ggplot(data = data, aes(x=distance, y=highly_rated)) +
  geom_smooth(method="loess", color="3a5e8cFF") +
  scale_y_continuous(expand = c(0.01,0.01),limits = c(0,1), breaks = seq(0,1,0.2), labels = scales::percent) +
  labs(x = "Distance",y = "Probability of Highly Rated") +
  theme_bw() +
  ggtitle("Probability of Highly Rated vs Distance") +
  theme(plot.title = element_text(size = 12), axis.title = element_text(size=8) )

```


```{r message=FALSE, warning=FALSE}
g2 <- ggplot(data = data) +
  geom_point(aes(x=pred_lpm, y=pred_probit, color="Probit"), size=0.5,  shape=16) +
  geom_point(aes(x=pred_lpm, y=pred_logit,  color="Logit"), size=0.5,  shape=16) +
  geom_line(aes(x=pred_lpm, y=pred_lpm,    color="45 Degree line"), size=0.5) +
  labs(x = "Predicted probability of Highly Rated (LPM)", y="Predicted probability")+
  scale_y_continuous(expand = c(0.00,0.0), limits = c(0,1), breaks = seq(0,1,0.2)) +
  scale_x_continuous(expand = c(0.00,0.0), limits = c(0,1), breaks = seq(0,1,0.2)) +
  scale_color_manual(name = "", values=c("#541352FF", "#3a5e8cFF","#10a53dFF")) +
  theme_bw() +
theme(legend.position=c(0.55,0.08),
        legend.direction = "horizontal",
        legend.text = element_text(size = 7))

```







